---
title: "M9/Final"
author: "Git In"
date: "12/11/2021"
output: html_document
bibliography: BIOL3140.bib

---
## Introduction
Trans-Gulf migrants (TGMs) travel long distances to forage and breed. These birds migrate from Central and South America every spring to more temperate regions between Texas and Florida. Their time of arrival to the breeding grounds have differing tradeoffs and can affect the fitness of the migrants. 

## Methods
In order to study the effect of local weather conditions on the arrival time of TGMs in Massachusetts, the meteorological data from Mobible AL, Charlotte NC, and Boston MA were acquired from NOAA’s National Climatic Data Center’s API to examine five species of birds (Hylocichla mustelina, Archilochus colubris, Seiurus aurocapilla, Progne subis, and Icterus galbula) from eBird.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgbif)
library(tidyverse)
library(MuMIn)
library(rnoaa)
library(data.table)
library(ggmap)
library(usmap)
library(rgdal)
library(magick)#for examples
library(cowplot)#for examples
library(lme4) #for linear mixed effect models
library(car) #for LME anova testing
library(data.table) #for frollmean function (and others)
```

```{r load data, cache=TRUE, include=FALSE}

dat <- readRDS("massbird1.data.RDS")
```

```{r just to visualize can remove later}
dat%>%
  group_by(year,species)%>%
  summarise(count=sum(individualCount,na.rm = T))%>%
  ggplot(aes(x=year,y=count,col=species))+geom_point()

```

```{r station info and weather, cache=TRUE, echo=FALSE}
options(noaakey = "lZNaAwTFydsOADUFQmqpKIIbVZejiypt")


sts <- c(
  "GHCND:USW00013894", #Mobible, AL 2k away about 10 days away @200 km/day
  "GHCND:USW00013881", #Charlotte, NC 1000 km away about 6 days away @200 km/day
  "GHCND:USW00014739" #Boston
)

sta.d <- bind_rows(
  lapply(sts,function(x) ncdc_stations(stationid = x)$data ) #use lapply to run through stations
  )%>%
  left_join(usmap_transform(.[,c("longitude","latitude")]))%>% #join transformation of lat/long for projection with usmap
  mutate(name=str_sub(name, -5,-4))%>%#simplify the name column, grab just the state
  mutate(migr.day=c(10,5,0))%>% #so we can look at wind speed 0, 5 or 10 days before arrive in boston
  separate(id,into = c("station.type","id"))%>%#need to cut station type out from station id number
        print()

plot_usmap(
  include = c(.northeast_region,.south_region,.east_north_central)
)+geom_point(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name),size=5)+geom_label(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name,label=name),size=5,nudge_x = 1e6*0.25)+theme(legend.position = "none")

#weather data from the staions
weather.d <- meteo_pull_monitors(sta.d$id,date_min = "2000-01-01")

#look at proportion of hylocichla mustelina that have arrived by each day and plot for each year
hm<- dat%>%
  filter(species=="Hylocichla mustelina")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)), j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))))%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)
hm%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)


hm.pred <- hm%>%
  group_by(species,year)%>%
 summarize(
   pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),#predict the logistic curve for each species
   j.day=min(j.day):max(j.day),
  )%>%
  left_join(hm%>%select(j.day,date)) ## add date back to tibble

hm%>%
  ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=hm.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)

hm.pred <- hm%>%
  group_by(year)%>%
 summarize(
   pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),#predict the logistic curve for each species
   j.day=min(j.day):max(j.day),
  )%>%
  left_join(hm%>%select(j.day,date)) ## add date back to tibble
```

```{r arrival proportions}

arrivals<- dat%>%
  group_by(species, year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)), j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))))%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)

arrivals.pred <- arrivals%>%
  group_by(species,year)%>%
 summarize(
   pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),#predict the logistic curve for each species
   j.day=min(j.day):max(j.day),
  )%>%
  left_join(arrivals%>%select(j.day,date)) ## add date back to tibble

arrivals%>%
  ggplot(aes(j.day,prop, col=species))+geom_point(aes=0.3)+geom_line(data=arrivals.pred,aes(x=j.day,y=pred),size=2)+facet_wrap(year~.)

```

```{r weather data}

arrival.date <-arrivals.pred%>%
  group_by(species, year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

weather.d <- weather.d%>%
  dplyr::mutate(year=as.integer(str_sub(date,1,4)), #add year
         date=as.Date(date))%>%
  group_by(year)%>% #group by year so we can compute julian day
  dplyr::mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), #add julian day
  date2=date,
  wdir.rad=(180-abs(wdf2-180))*pi/180, #radians so we can use a trig function to compute wind vector, scale degrees first to 180 scale to 2x pi and subtract from 180 (wind comes out of a direction)
  wvec=cos(wdir.rad)*-1*awnd # we want a negative value for positive value for 2x pi
  )%>% #store day in new column
  select(id,year,date2,j.day,tmin,tmax,wvec)%>% #select the rows we need
  left_join(sta.d%>%select(id,name,migr.day))%>% #add the station id info (ie. name)
  mutate(j.day=j.day+migr.day)#make j.day ahead of BOS according to the migration days away so we can join weather along path

arrival.weath <- arrival.date%>%
  left_join(weather.d)%>%
  left_join(arrivals%>%select(year,date,j.day))

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
         )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
arrivals.weath2 <- arrival.date%>%
  left_join(weather.wk)
```

```{r}
options(noaakey = "eOXSnLpzEZBYwjYJqQwwgBMMFLBbxGWz")
sts <- c(
  "GHCND:USW00013894", #Mobible, AL 2k away about 10 days away @200 km/day
  "GHCND:USW00013881", #Charlotte, NC 1000 km away about 6 days away @200 km/day
  "GHCND:USW00014739" #Boston
)
bos <- ncdc_stations(stationid = "GHCND:USW00014739")

print(bos)
```