---
title: "M9/Final"
author: "Git In"
date: "12/11/2021"
output: html_document
bibliography: BIOL3140.bib
---

## Introduction
Trans-Gulf migrants (TGMs) travel long distances to forage and breed. These birds migrate from Central and South America every spring to more temperate regions between Texas and Florida. Their time of arrival to the breeding grounds have differing tradeoffs and can affect the fitness of the migrants. Early arrivals correlate with higher chances of mating and breeding while later arrivals benefit from surplus food availiblility (@smith2005arrival). One study that studied migration patterns from 1828-2010, found a uniform trend of earlier arrivals correlated with higher temperatures from 1978 to 2010 (@kolavrova2017changes). 

In the face of climate change, populations of TGMs may decline if there is a failure to shift arrival dates in response to warmer temperatures (@both2006climate). 

With an increase in species conservation among the public in the recent decades, many online programs are available for researchers to access and study. The Cornell Laboratory of Ornithology and the National Audubon Society have established the program eBird, a compilation of over a hundred million bird observations from thousands of contributors.


## Methods
In order to study the effect of local weather conditions on the arrival time of TGMs in Massachusetts, the meteorological data from Mobile AL, Charlotte NC, and Boston MA were acquired from NOAA’s National Climatic Data Center’s API to examine five species of birds (*Hylocichla mustelina, Archilochus colubris, Seiurus aurocapilla, Progne subis*, and *Icterus galbula*) from eBird.

The arrival time for each species was considered to be the julian day when 25% of birds had arrived in a given year according to a logistic model. The birds were estimated to be in Mobile and Charlotte 10 and 5 days before the arrival in Boston, respectively. Using this, a rolling mean of temperature and wind data was taken over a two week period surrounding the dates that the birds were estimated to be passing through each location. 

Linear Mixed-Effect Modeling was used to evaluate the effect of multiple covariates and their individual and combined effect on arrival date. We ranked our nested models using the dredge() function in order to determine what was included in our best fit model. We then performed an ANOVA analysis on the best fit model to determine which variables had a significant effect on arrival date.

We also performed ANOVA analysis on maximum temperature, minimum temperature, and wind using a single day sample and two week average to determine the effect these variables have on arrival date. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgbif)
library(knitr)
library(tidyverse)
library(MuMIn)
library(rnoaa)
library(data.table)
library(ggmap)
library(usmap)
library(magick)#for examples
library(cowplot)#for examples
library(lme4) #for linear mixed effect models
library(car) #for LME anova testing
library(data.table) #for frollmean function (and others)
library(tidyverse) #always
library(nlme) #ditto
library(MuMIn)
library(knitr) #for knitting options
require(broom)
library(features)
```

```{r load data, cache=TRUE, include=FALSE}
dat <- readRDS("bird.data.RDS")
```

```{r station and weather data,cache=TRUE,echo=FALSE,warning=FALSE, error=FALSE,message=FALSE,fig.cap= "Figure 1: Map of the weather stations used along the migration route"}
options(noaakey = "lZNaAwTFydsOADUFQmqpKIIbVZejiypt")

sts <- c(
  "GHCND:USW00013894", #Mobible, AL 2k away about 10 days away @200 km/day
  "GHCND:USW00013881", #Charlotte, NC 1000 km away about 6 days away @200 km/day
  "GHCND:USW00014739" #Boston
)

sta.d <- bind_rows(
  lapply(sts,function(x) ncdc_stations(stationid = x)$data ) #use lapply to run through stations
  )%>%
  left_join(usmap_transform(.[,c("longitude","latitude")]))%>% #join transformation of lat/long for projection with usmap
  mutate(name=str_sub(name, -5,-4))%>%#simplify the name column, grab just the state
  mutate(migr.day=c(10,5,0))%>% #so we can look at wind speed 0, 5 or 10 days before arrive in boston
  separate(id,into = c("station.type","id"))#need to cut station type out from station id number

plot_usmap(
  include = c(.northeast_region,.south_region,.east_north_central)
)+geom_point(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name),size=5)+geom_label(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name,label=name),size=5,nudge_x = 1e6*0.25)+theme(legend.position = "none")

#weather data from the staions
weather.d <- meteo_pull_monitors(sta.d$id,date_min = "2000-01-01")

```

Weather data from three stations, Mobile AL, Charlotte NC, and Boston MA were used to estimate weather conditions along the migration route to Massachusetts. Weather parameters assessed include maximum temperature, minimum temperature, and the wind vector. 

```{r arrival proportions,echo=FALSE,warning=FALSE,error=FALSE, message=FALSE,fig.cap="Figure 2: logistic models representing the proportion of birds of each species arriving in Massachusetts each year."}
arrivals<- dat%>%
  group_by(species, year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)), j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))))%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)

arrivals.pred <- arrivals%>%
  group_by(species,year)%>%
 summarize(
   pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),#predict the logistic curve for each species
   j.day=min(j.day):max(j.day),
  )%>%
  left_join(arrivals%>%select(j.day,date)) ## add date back to tibble

arrivals%>%
  ggplot(aes(j.day,prop, col=species))+geom_point(aes=0.1)+geom_line(data=arrivals.pred,aes(x=j.day,y=pred),size=0.5)+facet_wrap(year~.)+geom_line(col="black",aes(y=0.25))

```

The proportion of birds arriving in Massachusetts in a given year was modelled using a logistic curve. 

```{r weather data, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

arrival.date <-arrivals.pred%>%
  group_by(species, year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

weather.d <- weather.d%>%
  dplyr::mutate(year=as.integer(str_sub(date,1,4)), #add year
         date=as.Date(date))%>%
  group_by(year)%>% #group by year so we can compute julian day
  dplyr::mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), #add julian day
  date2=date,
  wdir.rad=(180-abs(wdf2-180))*pi/180, #radians so we can use a trig function to compute wind vector, scale degrees first to 180 scale to 2x pi and subtract from 180 (wind comes out of a direction)
  wvec=cos(wdir.rad)*-1*awnd # we want a negative value for positive value for 2x pi
  )%>% #store day in new column
  select(id,year,date2,j.day,tmin,tmax,wvec)%>% #select the rows we need
  left_join(sta.d%>%select(id,name,migr.day))%>% #add the station id info (ie. name)
  mutate(j.day=j.day+migr.day)#make j.day ahead of BOS according to the migration days away so we can join weather along path

arrival.weath <- arrival.date%>%
  left_join(weather.d)%>%
  left_join(arrivals%>%select(year,date,j.day))

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
         )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
arrival.weath2 <- arrival.date%>%
  left_join(weather.wk)
```

Our ANOVA analysis revealed a significant relationship between minimum temperature as well as the interaction between maximum temperature and wind vector on a single day and the date of arrival.

```{r linear mixed-effect modeling, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),arrival.weath,na.action = "na.fail")

Anova(lmer) %>% tidy %>% knitr::kable(caption="ANOVA Analysis on Arrival Date and Single Day Sample weather data") #Anova from the car package
```

Our ANOVA analysis on the two-week average weather data revealed a significant relationship between the following variables and arrival date: minimum temperature, wind vector, and the interaction between minimum and maximum temperature.

```{r linear mixed-effect modeling2, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#0Mean two week weather preceding arrival
lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),arrival.weath2,na.action = "na.fail")

Anova(lmer2) %>% tidy %>% knitr::kable(caption="ANOVA Analysis on Arrival Date and Two-week average weather data")
```

Our analysis using the dredge() function returned NA results for the interactions models. Therefore, the best fit model was one that included minimum temperature, maximum temperature, and wind vector. This model was then analyzed using ANOVA and it was found that minimum temperature and wind vector were significant predictors of arrival date.


```{r linear mixed-effect modeling3, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
arr.aic <- dredge(lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)

arr.kb <- kable(arr.aic[1:4],caption = "Fit values for nested models of the most complicated lme model") 

arr.aic[1:4] %>% knitr::kable(caption="Fit values for nested models of the most complicated lme model")


best.lmer <-  lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),arrival.weath2,na.action = "na.fail")

Anova(best.lmer) %>% tidy %>% knitr::kable(caption="ANOVA Analysis on Best Fit Model")

```

## Discussion

Global warming has become a steady issue in the news, especially with conservationists. The trans-Gulf migrants are a specific group that have been at the center of studies because of their long distances traveled to forage and breed. Weather data was collected from Mobile, AL; Charlotte, NC; and Boston, MA to determine the weather conditions. It was found that the maximum temperature and the minimum temperature both increased. When the proportion of birds arriving early in Massachusetts was studied on a logistic curve, it was determined that more birds were arriving earlier than in previous years.

## Author Contributions

The introduction was written by Belinda Wang and Steven LaFever. Methods were written by Sarah Figueroa and Belinda Wang. Results were added by Sarah Figueroa. The discussion was written by Emma Oss.

## References

