---
title: "Project7_draft"
author: "GitIn"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #always
library(ape) #for evo analysis
library(nlme) #ditto
library(MuMIn) #for AIC
library(geiger) #for eco analysis
library(knitr) #for knitting options
```

### Introduction

### Methods

### Results

```{r data, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#read in mammal data and process
mammal <- read_csv("mammal.temp2.csv")
#add column with deltaT
mammal <- mammal %>% mutate(deltaT = abs(T.high - T.low))
#rename mass.g to mass 
mammal <- rename(mammal, mass = mass.g)
#log-transform data
mammal <- mammal %>% mutate_at(c('T.high', 'T.low','deltaT','mass'),log)
```
We were able to reproduce the scaling relationships of Mortola and Lanthier (2004) for Tmin, Tmax, and T delta. For T.high, the intercept estimate was 3.648 and the slope estimate was -0.00103. The model had a significant relationship between intercept and T.high but the slope was not signficant.
```{r t.high,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#create plot of T.high vs mass
T.high_plot <- mammal %>% ggplot(aes(mass,T.high))+geom_point()+geom_smooth(method="lm")+ggtitle("T high vs mass")+theme(plot.title=element_text(hjust=0.5))
print(T.high_plot)

#create linear model based on non-phylogenetically corrected data
T.high.reg <- lm(T.high ~ mass,mammal)
print(summary(T.high.reg))
```
For T low, the intercept estimate was 3.577 and the slope estimate was 0.00309. Both the intercept and slope (mass) varied significantly with respect to T low.
```{r t.low,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#create plot of T.low vs mass
T.low_plot <- mammal %>% ggplot(aes(mass,T.low))+geom_point()+geom_smooth(method="lm")+ggtitle("T low vs mass")+theme(plot.title=element_text(hjust=0.5))
print(T.low_plot)

#create linear model based on non-phylogenetically corrected data
T.low.reg <- lm(T.low ~ mass,mammal)
print(summary(T.low.reg))
```
For T delta, the intercept estimate was 1.071 and the slope estimate was -0.115. Both the intercept and slope (mass) varied significantly with T delta.
```{r deltaT.section,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#create plot of deltaT vs mass
deltaT_plot <- mammal %>% ggplot(aes(mass,deltaT))+geom_point()+geom_smooth(method="lm")+ggtitle("T delta vs mass")+theme(plot.title=element_text(hjust=0.5))
print(deltaT_plot)

#create linear model based on non-phylogenetically corrected data
deltaT.reg <- lm(deltaT ~ mass,mammal)
print(summary(deltaT.reg))
```
Significant phylogenetic signal was found for all traits by running Pagel's Lambda. 

```{r tree, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
library(ape)
library(phytools)

#read in phylogenetic tree
m.phy <- read.tree("mammal.tree.pruned.txt")
m.phy$tip.label <- gsub("(\\w+)_(\\w+)","\\1 \\2",m.phy$tip.label)

## Run Pagel's lambda for mass
mass.data <- mammal$mass
names(mass.data) <- mammal$species
mass.lambda <- phylosig(m.phy,mass.data,method="lambda",test=T)
mass.lambda.p <- mass.lambda$P
mass.lambda.lambda <- mass.lambda$lambda


## run Pagel's lambda for T.high
T.high.data <- mammal$T.high
names(T.high.data) <- mammal$species
T.high.lambda <- phylosig(m.phy,T.high.data,method="lambda",test=T)
T.high.lambda.p <- T.high.lambda$P
T.high.lambda.lambda <- T.high.lambda$lambda


## run Pagel's lambda for T.low
T.low.data <- mammal$T.low
names(T.low.data) <- mammal$species
T.low.lambda <- phylosig(m.phy,T.low.data,method="lambda",test=T)
T.low.lambda.p <- T.low.lambda$P
T.low.lambda.lambda <- T.low.lambda$lambda

## Run Pagel's lambda for deltaT
deltaT.data <- mammal$deltaT
names(deltaT.data) <- mammal$species
deltaT.lambda <- phylosig(m.phy,deltaT.data,method="lambda",test=T)
deltaT.lambda.p <- deltaT.lambda$P
deltaT.lambda.lambda <- deltaT.lambda$lambda

lambda_tibble <- tibble('Variable' = c('Mass', 'T high', 'T low', 'T delta'),'Lambda' = c(mass.lambda.lambda, T.high.lambda.lambda, T.low.lambda.lambda, deltaT.lambda.lambda), 'P-value' = c(mass.lambda.p, T.high.lambda.p, T.low.lambda.p, deltaT.lambda.p))
print(lambda_tibble)
#Run linear models on for each trait
```

Brownian Motion and Martins models were run on T low, T high, and T delta. The Martins models were better fits for each parameter.
```{r phy.models, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#Run Brownian Motion and OU Models
#T.low
T.low.BM.lm <- gls(T.low ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.high
T.high.BM.lm <- gls(T.high ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#deltaT
deltaT.BM.lm <- gls(deltaT ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.low
T.low.OU.lm <- gls(T.low ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.high
T.high.OU.lm <- gls(T.high ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

#deltaT
deltaT.OU.lm <- gls(deltaT ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

Bm.aic <- AICc(T.low.BM.lm,T.low.OU.lm,T.high.BM.lm,T.high.OU.lm,deltaT.BM.lm,deltaT.OU.lm)
Bm.aicw <- aicw(Bm.aic$AICc)
Bm.aicw <- Bm.aicw %>% add_column(Model = c('T low BM', 'T low OU', 'T high BM', 'T high OU', 'T delta BM', 'T delta OU'))
print(Bm.aicw)


```
The results from the phylogenetically-corrected models shows that there is not a significant relationship between T high and mass, and T low and mass, when phylogeny is taken into account.
```{r T.high.coef, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
T.low.coef <- coef(T.low.reg)
T.low.OU.coef <- coef(T.low.OU.lm)


T.high.coef <- coef(T.high.reg)
T.high.OU.coef <- coef(T.high.OU.lm)

deltaT.coef <- coef(deltaT.reg)
deltaT.OU.coef <- coef(deltaT.OU.lm)

#print summaries of all OU models since they are the better fits
print(anova(T.low.OU.lm))
print(anova(T.high.OU.lm))
print(anova(deltaT.OU.lm))
a <- anova(deltaT.OU.lm)
b <- a %>% select(c('p-value'))
c <- b[[1]][[1]]
print(c)

#create table with values for intercept in both models and their p values
i_tibble <- tibble('Variable' = c('T low', 'T low', 'T high', 'T high', 'T delta', 'T delta'), 'Model' = c('NPI', 'OU', 'NPI', 'OU', 'NPI', 'OU'), 'Value' = c(T.low.coef[1], T.low.OU.coef[1], T.high.coef[1], T.high.OU.coef[1], deltaT.coef[1], deltaT.OU.coef[1]))
print(i_tibble)

#create table with values for mass in both models and their p values
m_tibble <- tibble('Variable' = c('T low', 'T low', 'T high', 'T high', 'T delta', 'T delta'), 'Model' = c('NPI', 'OU', 'NPI', 'OU', 'NPI', 'OU'), 'Value' = c(T.low.coef[2], T.low.OU.coef[2], T.high.coef[2], T.high.OU.coef[2], deltaT.coef[2], deltaT.OU.coef[2]), 'P-value' = c())
print(m_tibble)

```
## Including Plots


## Discussion

## Author Contributions

## References