---
title: "Project7_draft"
author: "GitIn"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

Organismal biologists have a tendency to, in great error, forget to include phylogenetic structure in their research analyses [@Freckleton2002]. When discussing traits and how they vary in the context of another, it is important to include some phylogenetic data. This study aims to assess the scaling relationship of maximum and minimum daily body temperature between human and 52 species of non-human mammals [@Mortola2004].

### Methods

Data was collected from Mortola and Lanthier [@Mortola2004]. To consider phylogeny, a trimmed down phylogenetic tree of the taxa included in the study was used [@Upham2019]. Log-log regression models were made to compare Tb values as a function of mass. Secondly, to determine the phylogenetic signal of the four variables in the study, Pagel's lambda analysis was used.

### Results 

```{r echo=FALSE}
library(geiger)
library(tidyverse)

#read in mammal data and process
mammal <- read_csv("mammal.temp2.csv")
#add column with deltaT
mammal <- mammal %>% mutate(deltaT = abs(T.high - T.low))
#rename mass.g to mass 
mammal <- rename(mammal, mass = mass.g)
#log-transform data
mammal <- mammal %>% mutate_at(c('T.high', 'T.low','deltaT','mass'),log)



#create plot of T.high vs mass
T.high_plot <- mammal %>% ggplot(aes(mass,T.high))+geom_point()+geom_smooth(method="lm")
print(T.high_plot)


#create plot of T.low vs mass
T.low_plot <- mammal %>% ggplot(aes(mass,T.low))+geom_point()+geom_smooth(method="lm")
print(T.low_plot)


#create plot of deltaT vs mass
deltaT_plot <- mammal %>% ggplot(aes(mass,deltaT))+geom_point()+geom_smooth(method="lm")
print(deltaT_plot)
```

```{r tree, echo=FALSE}
library(ape)
library(phytools)

#read in phylogenetic tree
m.phy <- read.tree("mammal.tree.pruned.txt")
m.phy$tip.label <- gsub("(\\w+)_(\\w+)","\\1 \\2",m.phy$tip.label)
m.phy0<- rescale(m.phy, model = "lambda", 0)
m.phy5 <- rescale(m.phy, model = "lambda", 0.5)

par(mfcol = c(1, 3))
plot(m.phy)
plot(m.phy5)
plot(m.phy0)

## Run Pagel's lambda for mass
mass.data <- mammal$mass
names(mass.data) <- mammal$species
print(mass.data)
mass_lambda <- phylosig(m.phy,mass.data,method="lambda",test=T)
print(mass_lambda)

## run Pagel's lambda for T.high
T.high.data <- mammal$T.high
names(T.high.data) <- mammal$species
print(T.high.data)
T.high.lambda <- phylosig(m.phy,T.high.data,method="lambda",test=T)
print(T.high.lambda)


## run Pagel's lambda for T.low
T.low.data <- mammal$T.low
names(T.low.data) <- mammal$species
print(T.low.data)
T.low.lambda <- phylosig(m.phy,T.low.data,method="lambda",test=T)
print(T.low.lambda)


## Run Pagel's lambda for deltaT
deltaT.data <- mammal$deltaT
names(deltaT.data) <- mammal$species
print(deltaT.data)
deltaT_lambda <- phylosig(m.phy,deltaT.data,method="lambda",test=T)
print(deltaT_lambda)



```

### Discussion


### Author Contributions

### References