---
title: "M7_Draft_SF"
author: "Git In"
date: "10/25/2021"
output: html_document
bibliography: BIOL3140.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #always
library(ape) #for evo analysis
library(nlme) #ditto
library(MuMIn) #for AIC
library(geiger) #for eco analysis
library(knitr) #for knitting options
```

### Introduction
Organismal biologists have a tendency to, in great error, forgo including phylogenetic structure in their research analyses [@Freckleton2002]. When discussing traits and how they vary in the context of another, it is important to include some phylogenetic data. Species cannot be considered to be statistically independent since they are interrelated in a structured phylogeny  [@Felsenstein1985]. This isn't always possible depending on the availability of phylogenetic data, but with advances in evolutionary biology, this issue is becoming less prevalent. This study aims to assess the scaling relationship of maximum and minimum daily body temperature between human and 52 species of non-human mammals [@MortolaLanthier2004].Then, this relationship can be assessed in a phylogenetic context using phylogenetic least squares under Brownian and Ornstein-Uhlenbeck models. The phylogenetic signal is assessed using Pagel's lamba, a way to measure how much correlations in traits reflect their shared evolutionary history [@Pagel1999]. 

### Methods

Data was collected from Mortola and Lanthier [@MortolaLanthier2004]. To consider phylogeny, a trimmed down phylogenetic tree of the taxa included in the study was used [@Upham2019]. Log-log regression models were made to compare Tb values as a function of mass. Secondly, to determine the phylogenetic signal of the four variables in the study, Pagel's lambda analysis was used.

### Results

```{r data, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#read in mammal data and process
mammal <- read_csv("mammal.temp2.csv")
#add column with deltaT
mammal <- mammal %>% mutate(deltaT = abs(T.high - T.low))
#rename mass.g to mass 
mammal <- rename(mammal, mass = mass.g)
#log-transform data
mammal <- mammal %>% mutate_at(c('T.high', 'T.low','deltaT','mass'),log)
```
We were able to reproduce the scaling relationships of Mortola and Lanthier (2004) for Tmin, Tmax, and T delta. For T.high, the intercept estimate was 3.648 and the slope estimate was -0.00103. The model had a significant relationship between intercept and T.high but the slope was not signficant (Tables 3 & 4, Row 1).

```{r t.high,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#create plot of T.high vs mass
T.high_plot <- mammal %>% ggplot(aes(mass,T.high))+geom_point()+geom_smooth(method="lm")+ggtitle("T high vs mass")+theme(plot.title=element_text(hjust=0.5))
print(T.high_plot)

#create linear model based on non-phylogenetically corrected data
T.high.reg <- lm(T.high ~ mass,mammal)
```

Figure 1: Linear model of T high vs mass on non-phylogenetically corrected log-transformed data

For T low, the intercept estimate was 3.577 and the slope estimate was 0.00309. Both the intercept and slope (mass) varied significantly with respect to T low. (Tables 3 & 4, Row 3).

```{r t.low,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#create plot of T.low vs mass
T.low_plot <- mammal %>% ggplot(aes(mass,T.low))+geom_point()+geom_smooth(method="lm")+ggtitle("T low vs mass")+theme(plot.title=element_text(hjust=0.5))
print(T.low_plot)

#create linear model based on non-phylogenetically corrected data
T.low.reg <- lm(T.low ~ mass,mammal)
```

Figure 2: Linear model of T low vs mass on non-phylogenetically corrected log-transformed data

For T delta, the intercept estimate was 1.071 and the slope estimate was -0.115. Both the intercept and slope (mass) varied significantly with T delta. (Tables 3 & 4, Row 5).

```{r deltaT.section,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#create plot of deltaT vs mass
deltaT_plot <- mammal %>% ggplot(aes(mass,deltaT))+geom_point()+geom_smooth(method="lm")+ggtitle("T delta vs mass")+theme(plot.title=element_text(hjust=0.5))
print(deltaT_plot)

#create linear model based on non-phylogenetically corrected data
deltaT.reg <- lm(deltaT ~ mass,mammal)
```

Figure 3: Linear model of T delta vs mass on non-phylogenetically corrected log-transformed data

Significant phylogenetic signal was found for all traits by running Pagel's Lambda. 

Table 1: Pagel's Lambda Values

```{r tree, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
library(ape)
library(phytools)

#read in phylogenetic tree
m.phy <- read.tree("mammal.tree.pruned.txt")
m.phy$tip.label <- gsub("(\\w+)_(\\w+)","\\1 \\2",m.phy$tip.label)

## Run Pagel's lambda for mass
mass.data <- mammal$mass
names(mass.data) <- mammal$species
mass.lambda <- phylosig(m.phy,mass.data,method="lambda",test=T)
mass.lambda.p <- mass.lambda$P
mass.lambda.lambda <- mass.lambda$lambda


## run Pagel's lambda for T.high
T.high.data <- mammal$T.high
names(T.high.data) <- mammal$species
T.high.lambda <- phylosig(m.phy,T.high.data,method="lambda",test=T)
T.high.lambda.p <- T.high.lambda$P
T.high.lambda.lambda <- T.high.lambda$lambda


## run Pagel's lambda for T.low
T.low.data <- mammal$T.low
names(T.low.data) <- mammal$species
T.low.lambda <- phylosig(m.phy,T.low.data,method="lambda",test=T)
T.low.lambda.p <- T.low.lambda$P
T.low.lambda.lambda <- T.low.lambda$lambda

## Run Pagel's lambda for deltaT
deltaT.data <- mammal$deltaT
names(deltaT.data) <- mammal$species
deltaT.lambda <- phylosig(m.phy,deltaT.data,method="lambda",test=T)
deltaT.lambda.p <- deltaT.lambda$P
deltaT.lambda.lambda <- deltaT.lambda$lambda

lambda_tibble <- tibble('Variable' = c('Mass', 'T high', 'T low', 'T delta'),'Lambda' = c(mass.lambda.lambda, T.high.lambda.lambda, T.low.lambda.lambda, deltaT.lambda.lambda), 'P-value' = c(mass.lambda.p, T.high.lambda.p, T.low.lambda.p, deltaT.lambda.p))
print(lambda_tibble)
#Run linear models on for each trait
```

Brownian Motion (BM) and Ornstein-Uhlenbeck (OU) models were run on T low, T high, and T delta. The OU models were better fits for each parameter.


Table 2: AIC Scores of BM and OU Models

```{r phy.models, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#Run Brownian Motion and OU Models
#T.low
T.low.BM.lm <- gls(T.low ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.high
T.high.BM.lm <- gls(T.high ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#deltaT
deltaT.BM.lm <- gls(deltaT ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.low
T.low.OU.lm <- gls(T.low ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.high
T.high.OU.lm <- gls(T.high ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

#deltaT
deltaT.OU.lm <- gls(deltaT ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

Bm.aic <- AICc(T.low.BM.lm,T.low.OU.lm,T.high.BM.lm,T.high.OU.lm,deltaT.BM.lm,deltaT.OU.lm)
Bm.aicw <- aicw(Bm.aic$AICc)
Bm.aicw <- Bm.aicw %>% add_column(Model = c('T low BM', 'T low OU', 'T high BM', 'T high OU', 'T delta BM', 'T delta OU'))
print(Bm.aicw)


```

We created linear regressions for both the non-phylogentically-informed (NPI) and phylogenetically-informed (OU) models and calculated the estimated intercepts and slope coefficiencts. The table below shows the data for the intercepts. The values for intercepts were equal in the NPI and OU model for T low and T high and the intercepts were significant. For T delta, the intercepts were different and the relationship was significant for the NPI model but not the OU model.

Table 3: Intercept data for NPI vs OU models

```{r T.high.coef, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
T.low.coef <- coef(T.low.reg)
T.low.OU.coef <- coef(T.low.OU.lm)


T.high.coef <- coef(T.high.reg)
T.high.OU.coef <- coef(T.high.OU.lm)

deltaT.coef <- coef(deltaT.reg)
deltaT.OU.coef <- coef(deltaT.OU.lm)

#print summaries of all OU models since they are the better fits
la <- anova(T.low.OU.lm)[, c('p-value')]
ha <- anova(T.high.OU.lm)[, c('p-value')]
da <- anova(deltaT.OU.lm)[, c('p-value')]
la2 <- summary(T.low.reg)$coefficients[, c('Pr(>|t|)')]
ha2 <- summary(T.high.reg)$coefficients[, c('Pr(>|t|)')]
da2 <- summary(deltaT.reg)$coefficients[, c('Pr(>|t|)')]

#create table with values for intercept in both models and their p values
i_tibble <- tibble('Variable' = c('T low', 'T low', 'T high', 'T high', 'T delta', 'T delta'), 'Model' = c('NPI', 'OU', 'NPI', 'OU', 'NPI', 'OU'), 'Intercept' = c(T.low.coef[1], T.low.OU.coef[1], T.high.coef[1], T.high.OU.coef[1], deltaT.coef[1], deltaT.OU.coef[1]), 'P-value' = c(la2[[1]],la[[1]],ha2[[1]],ha[[1]],da2[[1]],da[[1]]), 'Signficant'=c('Yes','Yes','Yes','Yes','Yes','No'))

print(i_tibble)
```

The table below summarizes the slope coefficient data from the non-phylogentically-informed (NPI) and phylogenetically-informed (OU) models. The values for slope coefficient were different in the NPI and OU model for all three variables and for T low, the phylogenetically-informed model showed no significance where the NPI model showed significance. In other words, when you consider phylogeny, the significant correlation between T low and mass that Mortola and Lanthier found does not exist.


Table 4: Slope Coefficients for NPI vs OU Models

```{r mass_tibble, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#create table with values for mass in both models and their p values
m_tibble <- tibble('Variable' = c('T low', 'T low', 'T high', 'T high', 'T delta', 'T delta'), 'Model' = c('NPI', 'OU', 'NPI', 'OU', 'NPI', 'OU'), 'Slope' = c(T.low.coef[2], T.low.OU.coef[2], T.high.coef[2], T.high.OU.coef[2], deltaT.coef[2], deltaT.OU.coef[2]), 'P-value' = c(la2[[2]],la[[2]],ha2[[2]],ha[[2]],da2[[2]],da[[2]]), 'Signficant'=c('Yes','No','No','No','Yes','Yes'))
print(m_tibble)

```

## Discussion

## Author Contributions

## References