---
title: "Project 7"
author: "Git In"
date: "11/02/2021"
output: html_document
bibliography: BIOL3140.bib
---

<!-- for comments -->
<style>
div.comment pre { background-color:#ffbbbb; }
div.comment pre.r { background-color:#ffbbbb; }
div.comment {color: red;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #always
library(ape) #for evo analysis
library(nlme) #ditto
library(MuMIn) #for AIC
library(geiger) #for eco analysis
library(knitr) #for knitting options
require(broom)
```

### Introduction
Organismal biologists have a tendency to, in great error, forgo including phylogenetic structure in their research analyses [@Freckleton2002]. When discussing traits and how they vary in the context of another, it is important to include some phylogenetic data. Species cannot be considered to be statistically independent since they are interrelated in a structured phylogeny  [@Felsenstein1985]. This isn't always possible depending on the availability of phylogenetic data, but with advances in evolutionary biology, this issue is becoming less prevalent. This study aims to assess the scaling relationship of maximum and minimum daily body temperature between human and 52 species of non-human mammals [@MortolaLanthier2004].Then, this relationship can be assessed in a phylogenetic context using phylogenetic least squares under Brownian and Ornstein-Uhlenbeck models. The phylogenetic signal is assessed using Pagel's lamba, a way to measure how much correlations in traits reflect their shared evolutionary history [@Pagel1999]. 


### Methods

Data was collected from Mortola and Lanthier [@MortolaLanthier2004]. To consider phylogeny, a trimmed down phylogenetic tree of the taxa included in the study was used [@Upham2019]. Log-log regression models were made to compare Tb values (T low, T high, and T delta) as a function of mass. We also reproduced the graphs from the [@MortolaLanthier2004] paper.

Secondly, to determine the phylogenetic signal of the four variables in the study, Pagel's lambda analysis was used. Pagel's lambda is a method that mutes phylogenetic signal by lengthening branch tips to determine how much of a trait is driven by phylogeny. A lambda value of zero means there is no phylogenetic signal and a lambda value of one means that the data conforms to a Brownian Motion Model.

After running the Pagel's lambda analysis. Brownian Motion (BM) and Ornstein-Uhlenbeck (OU) models were run on T low, T high, and T delta to see which phylogenetically-informed model fit the data better. The OU models were better fits for each parameter. Therefore, the OU models were used in further tests. We created linear regressions for both the non-phylogentically-informed (NPI) and phylogenetically-informed (OU) models and calculated the estimated intercepts and slope coefficiencts (Tables 3&4).


### Results

```{r data, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}

#read in mammal data and process
mammal <- read_csv("mammal.temp2.csv")
#add column with deltaT
mammal <- mammal %>% mutate(deltaT = abs(T.high - T.low))
#rename mass.g to mass 
mammal <- rename(mammal, mass = mass.g)
#log-transform data
mammal <- mammal %>% mutate_at(c('T.high', 'T.low','deltaT','mass'),log)
```
We were able to reproduce the scaling relationships of Mortola and Lanthier (2004) for Tmin, Tmax, and T delta. 

```{r t.low,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE, fig.cap="Figure 1: Graph of T low vs mass"}

#create plot of T.low vs mass
T.low_plot <- mammal %>% ggplot(aes(mass,T.low))+geom_point()+geom_smooth(method="lm")+ggtitle("T low vs mass")+theme(plot.title=element_text(hjust=0.5))
print(T.low_plot)

#create linear model based on non-phylogenetically corrected data
T.low.reg <- lm(T.low ~ mass,mammal)
```

```{r t.high,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE, fig.cap="Figure 2: Graph of T high vs mass"}

#create plot of T.high vs mass
T.high_plot <- mammal %>% ggplot(aes(mass,T.high))+geom_point()+geom_smooth(method="lm")+ggtitle("T high vs mass")+theme(plot.title=element_text(hjust=0.5))
print(T.high_plot)

#create linear model based on non-phylogenetically corrected data
T.high.reg <- lm(T.high ~ mass,mammal)
```


```{r deltaT.section,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE,fig.cap="Figure 3: Graph of T delta vs mass"}

#create plot of deltaT vs mass
deltaT_plot <- mammal %>% ggplot(aes(mass,deltaT))+geom_point()+geom_smooth(method="lm")+ggtitle("T delta vs mass")+theme(plot.title=element_text(hjust=0.5))
print(deltaT_plot)

#create linear model based on non-phylogenetically corrected data
deltaT.reg <- lm(deltaT ~ mass,mammal)
```




Significant phylogenetic signal was found for all traits by running Pagel's Lambda and phylogenetic signal was high (close to 1).


```{r tree, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
library(ape)
library(phytools)

#read in phylogenetic tree
m.phy <- read.tree("mammal.tree.pruned.txt")
m.phy$tip.label <- gsub("(\\w+)_(\\w+)","\\1 \\2",m.phy$tip.label)

## Run Pagel's lambda for mass
mass.data <- mammal$mass
names(mass.data) <- mammal$species
mass.lambda <- phylosig(m.phy,mass.data,method="lambda",test=T)
mass.lambda.p <- mass.lambda$P
mass.lambda.lambda <- mass.lambda$lambda


## run Pagel's lambda for T.high
T.high.data <- mammal$T.high
names(T.high.data) <- mammal$species
T.high.lambda <- phylosig(m.phy,T.high.data,method="lambda",test=T)
T.high.lambda.p <- T.high.lambda$P
T.high.lambda.lambda <- T.high.lambda$lambda


## run Pagel's lambda for T.low
T.low.data <- mammal$T.low
names(T.low.data) <- mammal$species
T.low.lambda <- phylosig(m.phy,T.low.data,method="lambda",test=T)
T.low.lambda.p <- T.low.lambda$P
T.low.lambda.lambda <- T.low.lambda$lambda

## Run Pagel's lambda for deltaT
deltaT.data <- mammal$deltaT
names(deltaT.data) <- mammal$species
deltaT.lambda <- phylosig(m.phy,deltaT.data,method="lambda",test=T)
deltaT.lambda.p <- deltaT.lambda$P
deltaT.lambda.lambda <- deltaT.lambda$lambda

lambda_tibble <- tibble('Variable' = c('Mass', 'T high', 'T low', 'T delta'),'Lambda' = c(mass.lambda.lambda, T.high.lambda.lambda, T.low.lambda.lambda, deltaT.lambda.lambda), 'P-value' = c(mass.lambda.p, T.high.lambda.p, T.low.lambda.p, deltaT.lambda.p))
lambda_tibble_df <- as.data.frame(lambda_tibble)
lambda_tibble_df%>%
  knitr::kable(caption="Table 1: Pagel's Lambda Values")
#Run linear models on for each trait
```

AIC scores were lower for the OU model for every trait.

```{r phy.models, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#Run Brownian Motion and OU Models
#T.low
T.low.BM.lm <- gls(T.low ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.high
T.high.BM.lm <- gls(T.high ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#deltaT
deltaT.BM.lm <- gls(deltaT ~ mass, correlation = corBrownian(1,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.low
T.low.OU.lm <- gls(T.low ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

#T.high
T.high.OU.lm <- gls(T.high ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

#deltaT
deltaT.OU.lm <- gls(deltaT ~ mass, correlation = corMartins(0.00000001,phy = m.phy,form=~species),data = mammal, method = "ML")

Bm.aic <- AICc(T.low.BM.lm,T.low.OU.lm,T.high.BM.lm,T.high.OU.lm,deltaT.BM.lm,deltaT.OU.lm)
Bm.aicw <- aicw(Bm.aic$AICc)
Bm.aicw <- Bm.aicw %>% add_column(Model = c('T low BM', 'T low OU', 'T high BM', 'T high OU', 'T delta BM', 'T delta OU'))
Bm.aicw%>%dplyr::mutate_if(is.numeric, funs(as.character(signif(., 4))))%>%
  knitr::kable(caption="Table 2: AIC Scores of BM and OU Models")


```

<div class="comment">
But we wanted to assess the fit of the phylo vs non-phylo models for each variable separately.
</div>

The values for intercept and slope coefficients from the graphs from [@MortolaLanthier2004] and the phylogenetically-informed OU models are below. The intercept varied signficantly with every variable for the NPI data. The intercept was significant for T low and T high for the OU data, but not for T delta. The values for intercepts were nearly equal in the NPI and OU models for T low and T high but were different for T delta.

```{r T.high.coef, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
T.low.coef <- coef(T.low.reg)
T.low.OU.coef <- coef(T.low.OU.lm)


T.high.coef <- coef(T.high.reg)
T.high.OU.coef <- coef(T.high.OU.lm)

deltaT.coef <- coef(deltaT.reg)
deltaT.OU.coef <- coef(deltaT.OU.lm)

#print summaries of all OU models since they are the better fits
la <- anova(T.low.OU.lm)[, c('p-value')]
ha <- anova(T.high.OU.lm)[, c('p-value')]
da <- anova(deltaT.OU.lm)[, c('p-value')]
la2 <- summary(T.low.reg)$coefficients[, c('Pr(>|t|)')]
ha2 <- summary(T.high.reg)$coefficients[, c('Pr(>|t|)')]
da2 <- summary(deltaT.reg)$coefficients[, c('Pr(>|t|)')]

#create table with values for intercept in both models and their p values
i_tibble <- tibble('Variable' = c('T low', 'T low', 'T high', 'T high', 'T delta', 'T delta'), 'Model' = c('NPI', 'OU', 'NPI', 'OU', 'NPI', 'OU'), 'Intercept' = c(T.low.coef[1], T.low.OU.coef[1], T.high.coef[1], T.high.OU.coef[1], deltaT.coef[1], deltaT.OU.coef[1]), 'P-value' = c(la2[[1]],la[[1]],ha2[[1]],ha[[1]],da2[[1]],da[[1]]), 'Signficant'=c('Yes','Yes','Yes','Yes','Yes','No'))
i_tibble_df <- as.data.frame(i_tibble)
i_tibble_df%>%
  knitr::kable(caption="Table 3: Intercept data for NPI vs OU models")
```

<div class="comment">
Really like how you did this! Presents the comparisons clearly!
</div>

The table below summarizes the slope coefficient data from the non-phylogentically-informed (NPI) and phylogenetically-informed (OU) models. The values for slope coefficient were different in the NPI and OU model for all three variables and for T low, the phylogenetically-informed model showed no significance where the NPI model showed significance. In other words, when you consider phylogeny, the significant correlation between T low and mass that Mortola and Lanthier found does not exist.

<div class="comment">
And Voila!!!
</div>
```{r mass_tibble, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#create table with values for mass in both models and their p values
m_tibble <- tibble('Variable' = c('T low', 'T low', 'T high', 'T high', 'T delta', 'T delta'), 'Model' = c('NPI', 'OU', 'NPI', 'OU', 'NPI', 'OU'), 'Slope' = c(T.low.coef[2], T.low.OU.coef[2], T.high.coef[2], T.high.OU.coef[2], deltaT.coef[2], deltaT.OU.coef[2]), 'P-value' = c(la2[[2]],la[[2]],ha2[[2]],ha[[2]],da2[[2]],da[[2]]), 'Signficant'=c('Yes','No','No','No','Yes','Yes'))
m_tibble_df <- as.data.frame(m_tibble)
m_tibble_df%>%
  knitr::kable(caption="Table 4: Slope Coefficients for NPI vs OU Models")

```


## Discussion
The analysis showed that the scaling relationships of the previous Mortola and Lanthier (2004) study for Tmin, Tmax, and T delta were able to be reproduced using log-log regression models. The four variables, mass, Tmin, Tmax, and TΔ, showed significant phylogenetic signals. A study from [@clarke2008scaling] also showed significant scaling of body temperature with mass in Marsupialia, Ferae, and Ungulata. Running Pagel's Lambda, the signal for all of them was high and close to 1. 
For all of the traits, the AIC scores were lower for the OU model versus the BM model, so it fits best. However, when running the models, the intercepts of Tmin and Tmax were significant for the OU data. TΔ was not found to be significant. Tmin and Tmax intercept values were almost equal in the NPI and OU models, but the TΔ value was different. A very important result we found is that T low does not depend significantly on mass when phylogeny is taken into account. This an important result because it disputes the results in [@MortolaLanthier2004] and demonstrates that they may have erred when they did not take phylogeny into account and treated species as independent samples.


## Author Contributions

The introduction was written by Sarah Figueroa and Emma Oss. Coding chunks and results were added by Steven LaFever. References were added by Sarah Figueroa. Methods were written by Steven LaFever and Belinda Wang. The discussion was written by Belinda Wang and Emma Oss.

## References
